---
title: "Exercise 2"
author: "EMT Learns R"
output:
  pdf_document: default
---


1. Add a YAML header to this Exercise 2 Starter .Rmd file. 


2. Install the following packages by putting this code in your console: install.packages("packagename") for each package. I want you to go ahead and install: tidyverse, haven, and readxl. After installed, load them using the library command here in this first code chunk.
```{r}

```


3. Locate the working directory for this .Rmd file.  
```{r}

```


4. Import the excel file named ("naloxonecleaned_allobservations_todate") from the M Drive (*Project folder 322-State Opioid Response (SOR) > 11 - Quarterly and Annual Reports > Data > Naloxone Inventory Data*). You will need to use the function read_excel instead of load (for more info about readxl, see here: https://readxl.tidyverse.org/). Also, make sure you assign a name to the dataset when you read it in by specifying a name (of your choosing) before a backwards arrow. See my example commented out in the code chunk below.  
```{r}
#Change what's needed in the code below to make this run. In my example, I'm calling the data "NALOXONE_DATA" when I import it into R. You will need to change everything that's in all caps to run this code (and un-comment it out so that R interprets it as code).

#NALOXONE_DATA<-read_excel("PATH/FILENAME.xlsx")
```


5. How many observations are in this dataset? How many variables are in this dataset? Where did you find that information in R? 



6. Knit this completed file to html and upload that to the Exercises folder (be sure to name this .Rmd file as follows, before you knit: FirstName_Exercise2.Rmd)



7. Knit this completed file to pdf and upload that to the Exercises folder. 



*Troubleshooting for knitting errors can be found on the EMTlearnsR webpage under the second >> under General R Resources (the slide deck). Slides toward the end give you Mac- or PC-specific troubleshooting steps when having trouble knitting. If you've tried everything in here (and google isn't helping either), email me a screen shot of the error you're still getting and what you've tried so far. The first knit is the hardest :) so don't lose heart if it takes some time to iron out kinks. 
