---
title: "Exercise 3"
author: "EMT Learns R"
output:
  pdf_document: default
---

Don't forget to add your YAML header to the top of this .Rmd file before you begin the exercise. 

1. You will need the following packages for this exercise: `readxl`, `haven`, and `tidyverse`. Load these in your first code chunk below.
```{r}

```

2. Import the excel dataset named ("naloxonecleaned_allobservations_todate") from the M Drive (*Project folder 322-State Opioid Response (SOR) > 11 - Quarterly and Annual Reports > Data > Naloxone Inventory Data*)
```{r}

```

3. Examine this dataset. 
```{r}

```
3a. What is the range of the number of units distributed (i.e., the minimum and the maximum)?      

3b. What is the range of distribution dates reflected in this dataset (i.e., the earliest and latest distributions)? 

4. Let's say we're just interested in examining the last quarter (Q4) of fiscal year 2021 (FY21). Create a new dataset (remember - give it a NEW name) that only includes FY21_Q4 observations. 
```{r}

```
4a. How many observations do we have for FY21_Q4? 


5. Next, create a count of total units distributed by region in FY21_Q4. 
```{r}

```
5a. Which region distributed the highest number of naloxone units in FY21_Q4? 


6. Plot the regional distribution of naloxone units for FY21_Q4. Be sure your plot is labeled appropriately and that everything is readable/understandable. 
```{r}

```

BONUS:
Which county received the highest number of naloxone units in FY2021 (the whole fiscal year, not just the last quarter)? 
