---
Title: "EMT Learns R - Session 5"
Date: "02-07-2022"
---

Session Plan:
1. Current projects in R? What have you been using R for? 
2. Upcoming plans in R. Future sessions will focus on workshopping current code, so you should plan to bring an EMT project to these sessions.
3. Review of what we've gone through so far and housekeeping (e.g., keep R and packages updated). 
4. Some new code - joining/merging files

```{r} 
library(haven)
library(readxl)
library(tidyverse)
```

Reading data in to R 
```{r}
exceldata<-read_excel("FILEPATH/FILENAME.xlsx")
csvdata<-read_csv("FILEPATH/FILENAME.csv")
statadata<-read_dta("FILEPATH/FILENAME.dta")
spssdata<-read_spss("FILEPATH/FILENAME.sav")
sasdata<-read_sas("FILEPATH/FILENAME.sas7bdat")
```

Descriptive/Exploratory Stats
```{r}
summary(DATASET) #5-number summary for dataset or continuous variables
table(DATASET$VARIABLE) #frequency table for categorical data
mean(DATASET$VARIABLE) #mean for continuous variable

#conditional means (mean of continuous variable by levels of categorical var)
DATASET%>%
    group_by(VARIABLE1)%>%
    summarize(NEWVARNAME=mean(VARIABLE2))
```

"Tidying"/Manipulating Data
```{r}
NEWVARNAME<-NEWVARNAME%>%
    select(VARIABLE)%>% #keep only the chosen variable/column
    filter(VARIABLE>10) #keep only those observations/rows where VARIABLE > 10
```

Housekeeping:
Be sure to check for package updates (at least once per month or so) and then Rstudio/R updates as well (HELP > CHECK FOR UPDATES in Rstudio). R updates less frequently; so maybe check this 1-2 times per year. You can see your R version at the top of your Console below.

NEW CODE: Merging/joining data files
For many of our projects, we need to pull data from multiple sources. R makes this easy using the set of "join" functions. I'm going to walk you through an example using TN Together Student Survey data. I have the full data file and want to add in response rates to the master file.
```{r include=F}
#Read in each data file
analysis<-read_dta("M:/334 - TN Together/000 - TN Together 2020-21 Administration/Analysis/output/state_analyticaldataset.dta")

resprate<-read_dta("M:/334 - TN Together/000 - TN Together 2020-21 Administration/Analysis/output/responserates.dta")

analysis_wrr<-left_join(analysis, resprate)
```
While I most frequently use left_join() there are also full_join() and right_join() functions. 

NEW CODE: Appending data files
Sometimes, we don't want to merge datasets, but instead we want to append them. This is when you have the same data on multiple sites, for example, and just want it all in one datafile. As usual in R there are multiple ways to do this, but my preferred is the bind_rows() function (note: there is also a bind_cols() function that does the same for columns).

The example below takes ROPS contact information that is currently saved on two separate excel sheets and binds them together (so that I have one file with all current/previous ROPS). 
```{r}
#ROPS# and Regions (2 sheets)
rops1<-read_excel("M:/347 - State Opioid Response (SOR II) Grant/Contact Information/ROPS - Contact Information.xlsx")

rops2<-read_excel("M:/347 - State Opioid Response (SOR II) Grant/Contact Information/ROPS - Contact Information.xlsx", sheet="Previous_ROPS")

ropsfull<-bind_rows(rops1, rops2)
```


